---
output: html_document
header-includes:
   - \usepackage{amsmath}
   - \setlength{\mathindent}{0pt}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 20)
```

<br>

## How to Calculate a Goal Score
***


![](https://docs.google.com/drawings/d/e/2PACX-1vRN8pFrclhIYIPk7wBzI1TSAkycNzIK7KUKmqcMtej8KBPnsm4RbhjIsGwUPrmJ-7aRhmdlBscCMGUe/pub?w=1658&h=584)

***

Each of the [ten goals](https://www.nature.com/articles/nature11397) have their own unique methodology. [Methodologies](https://rawgit.com/OHI-Science/ohi-global/draft/global_supplement/Supplement.html#5_goal_models_and_data) are built around the types of available data, appropriate for the analysis -- consistently quantified across the spatial scale of the assessment and quality-controlled -- and based on the best scientific understanding of the subject per expert input and peer review. Each [goal score calculation](https://github.com/OHI-Science/ohi-global/blob/draft/eez/conf/functions.R) takes a few different [data layers](http://ohi-science.org/ohi-global/layers_table.html) that inform present status and likely near-term future status, which is calculated from trend, pressure, and resilience information. To learn about how each goal is calculated click below. For more information about each goal visit [ohi-science.org](http://ohi-science.org/goals/).


***

```{r, echo = FALSE}
views <- c("goals_calc/ao_calc.png", "goals_calc/hab_calc.png", "goals_calc/spp_calc.png", "goals_calc/cp_calc.png",
           "goals_calc/cs_calc.png", "goals_calc/cw_calc.png", "goals_calc/fis_calc.png", "goals_calc/mar_calc.png",
           "goals_calc/eco_calc.png", "goals_calc/liv_calc.png", "goals_calc/np_calc.png", "goals_calc/ico_calc.png",
           "goals_calc/lsp_calc.png", "goals_calc/tr_calc.png")

names(views) <- c("Artisanal Fishing Opportunity", "Biodiversity: Habitat", "Biodiversity: Species", 
                  "Coastal Protection", "Carbon Storage", "Clean Waters", "Food Provision: Fisheries", "Food Provision: Mariculture",
                  "Livelihoods and Economies: Economies", "Livelihoods and Economies: Livelihoods", 
                  "Natural Products", "Sense of Place: Iconic Species", "Sense of Place: Lasting Special Places", "Tourism and Recreation")

bsselectR::bsselect(views, type = "img", show_tick = TRUE, height = "10")
```

***